### 2017-11-08 doc/figures/asgsr_scratch.tex
changes to figures

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Erin Schmidt}
\begin{document}
\subsection*{Playing with Scales}
In the inertial limit, and with $y \ll L$,
\[ m y'' + q \mathbf{E}_0 = 0 \hspace{2 mm} \mbox{subject to } \hspace{2 mm} y(0) = 0, \hspace{2 mm} y'(0) = U_0 \] 
\[ t_0 \sim \frac{m U_0}{q E_0}, \hspace{2 mm} y_0 \sim \frac{m U_0^2}{q E_0}\]
\[\frac{y_0}{L} \sim \frac{m U_0^2}{q E_0 L} \equiv {\mathbb{E}\mbox{u}}_e \equiv \frac{\mbox{inertia}}{\mbox{electrostatic force}}\]

\newpage
\subsection*{Parameter Estimation}
We find the parameters $\mathbf{x}$ that solve the inverse problem $G(\mathbf{x}) = D$, using a direct search method (\emph{Nelder-Mead}). 
\begin{equation*}
\mbox{min} \hspace{2 mm} \chi^2 = \mbox{min} \hspace{2 mm} \sum^n_{i=1} \frac{\left({yD(\mathbf{x})}_i - yG(\mathbf{x})_i \right)^2}{yG(\mathbf{x})_i}
\end{equation*}
\[\mbox{where} \hspace{2 mm} yG(\mathbf{x}) \hspace{2 mm} \mbox{is a solution of} \hspace{2 mm} m y'' = \frac{1}{2} \rho C_D A_d {y'}^2 + q E(y) + \frac{1}{4} \frac{K q^2}{y^2} + \frac{1}{2} {E(y)}^2 \nabla \epsilon\]
\begin{eqnarray*} \mbox{} \hspace{2 mm} \begin{split} \mathbf{x} = \left\{ \begin{array}{ll}      & q\\
		  &	V_d\\
          & \sigma 
          \end{array} \right. 
          \end{split} \hspace{2 mm} \mbox{subject to} \hspace{2 mm} \begin{split}
          g = \left\{ \begin{array}{ll}
           V_d &\pm \hspace{2 mm} u_{exp}\\
      	   \sigma &\pm  \hspace{2 mm} u_{exp}\\
      	   y_0 &\pm \hspace{2 mm} u_{exp}\\
      	   t_0 &\pm \hspace{2 mm} u_{exp}\\
          \end{array} \right. 
          \end{split}
\end{eqnarray*}
\end{document}